# Mouse Weights

### Overview
Extreme weight loss and loss of muscle mass, also known as cachexia, typically presents along side chronic inflammatory illnesses like Tuberculosis disease @ [baazim2022interplay]. We now recognize that cachexia is part of a systemic response to inflammation, and has been linked to upregulation of pro-inflammatory cytokines such as TNF, IL-6, and IFNg in humans @ [baazim2022interplay]. Additionally, studies support the role of key immune cell populations such as CD8+ T-cells that, when depleted, counteract muscle and fat deterioration  @[baazim2019cd8],  and suggest that CD8+ T-cells may metabolically reprogram adipose tissue.

In recognition of cachexia related illnesses and diseases, we tracked the progression of weight loss over the course of this study, as is done with many TB-mouse studies @ [smith2022host], @ [segueni2016controlled]. These data is also useful when correlating to CFU count as well as expression of cytokines and other biological markers @ [smith2022host]. Here, mice are weighed in grams weekly to monitor clinical status as TB patients frequently display weight loss as clinical symptom associated with disease progression. 

The following contains information about how the data was collected, organized, and curated for analysis in RStudio.

### Parameters
Weights are recorded in an excel worksheet. 

Column titles are as follows: who_collected	date_collected	sex	dob	notch_id	mouse_number	weight	unit	cage_number	group	notes

Groups included are: bcg, saline, bcg+id93, saline+id93, saline+noMtb

The notes column contains information regarding clinical observations. 

good reference: https://elifesciences.org/articles/74419#s4

```{r}
library(readxl)
library(tidyverse)
```

### Read in data

Data is stored in one excel sheet, each week is one sheet named as the date -> return vector for each sheet name 

```{r include=FALSE}
weight_data <- read_xlsx("DATA/body_weights_measurement.xlsx")
excel_sheets("DATA/body_weights_measurement.xlsx") #lists names of each sheet

multiplesheets <- function(fname) {
   
  # getting info about all excel sheets
  sheets <- readxl::excel_sheets(fname)
  tibble <- lapply(sheets, function(x) readxl::read_excel(fname, sheet = x))
  data_frame <- lapply(tibble, as.data.frame)
    
  # assigning names to data frames
  names(data_frame) <- sheets
    
  # print data frame
  print(data_frame)
}
  
# specifying the path name
path <- "DATA/body_weights_measurement.xlsx"
multiplesheets(path)
```

### Can also use rio to read in the data, more streamlined
```{r include=FALSE}
library(rio)
# specifying the path name
path <- "DATA/body_weights_measurement.xlsx"
  
# reading data from all sheets
data <- import_list(path)
  
# print data
print (data)
```


### Clean data
```{r}
dataset <- data %>%
 select("sex", "mouse_number", "weight", "cage_number", "group")

```


```{r}

# combining columns mouse_number and cage_number

dataset$mouse_id <- paste(dataset$mouse_number, "-", dataset$cage_number)

```


### Body weight over time graph and statistics
```{r}

```

### Weight loss over time graph and statistics 
```{r}

```

### Weight vs CFU
```{r}

```

### Weight vs ELISA results 
```{r}

```

### Weight vs lesion burden
```{r}

```

