# Mouse Weights

For example, @baazim2022interplay said ... 

Mice are weighed in grams weekly to monitor clinical status as TB patients frequently display weight loss as clinical symptom associated with disease progression.

Weights are recorded in an excel worksheet. 

Column titles are as follows: who_collected	date_collected	sex	dob	notch_id	mouse_number	weight	unit	cage_number	group	notes

Groups included are: bcg, saline, bcg+id93, saline+id93, saline+noMtb

The notes column contains information regarding clinical observations. 

good reference: https://elifesciences.org/articles/74419#s4

```{r}
library(readxl)
library(tidyverse)
```

## Read in data

# Data is stored in one excel sheet, each week is one sheet named as the date -> return vector for each sheet name 

```{r include=FALSE}
weight_data <- read_xlsx("DATA/body_weights.xlsx")
excel_sheets("DATA/body_weights.xlsx") #lists names of each sheet

multiplesheets <- function(fname) {
   
  # getting info about all excel sheets
  sheets <- readxl::excel_sheets(fname)
  tibble <- lapply(sheets, function(x) readxl::read_excel(fname, sheet = x))
  data_frame <- lapply(tibble, as.data.frame)
    
  # assigning names to data frames
  names(data_frame) <- sheets
    
  # print data frame
  print(data_frame)
}
  
# specifying the path name
path <- "DATA/body_weights.xlsx"
multiplesheets(path)
```

# can also use rio to read in the data, more streamlined
```{r include=FALSE}
library(rio)
# specifying the path name
path <- "DATA/body_weights.xlsx"
  
# reading data from all sheets
data <- import_list(path)
  
# print data
print (data)
```


## Clean data
```{r}
data$before_vaccination %>%
 select("sex", "notch_id",  "weight", "cage_number", "group") #"mouse_number" it couldn't find this for some reason 
```

## Summary statistics
```{r}

```


# Body weight over time graph and statistics
```{r}

```

# Weight loss over time graph and statistics 
```{r}

```

# Weight vs CFU
```{r}

```

# Weight vs ELISA results 
```{r}

```

# Weight vs lesion burden
```{r}

```

