---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Flow cytometry

Flow cytometry data can be quantified in many different ways and with different techniques. For the purpose of these data analyses, manual gating has been achieved in FlowJo and cell frequencies and populations exported as a `.csv` file. This `.csv` file is the primary input for this R pipeline which aims to output box plots for each gated cell population. 

This example data set is from an innate response study whcih investigated the immune response in the lungs during the first 28 days of infection.

----------------------------------------------------------------------------------

Immune cells are very diverse, and the make-up of immune cells within a sample
can provide important insights on immune processes based on measures of this
composition. Immune cells can be categorized into large groups (e.g., T cells, B
cells, macrophages, dendritic cells). They can also be characterized into
different populations within these large groups, based on things like activation
and differentiation [?] within the group [?] [@maecker2012standardizing] (e.g.,
T cells can be divided into naive T cells versus memory T cells, helper T cells
versus cytotoxic T cells, [others] [?]). This process of characterizing the
immune cells in a sample is called immunophenotyping.

To make these classifications, flow cytometry uses a pretty clever mix of 
physics and biology. First, it starts by leveraging the biological knowledge
that antibodies can have a very specific affinity for a certain protein. 
This means that you can find a set of antibodies that will target and 
stick specifically to certain proteins. 

Flow cytometry starts by creating a panel of up to [x] protein markers, focusing
on proteins that can help in identifying a specific cell type. Typically, the 
panel will include several proteins that are "CD" proteins (the "CD" stands for
cluster of differentiation or classification determinant). These are proteins
that show up on the surface of immune cells, with specific CD proteins common 
to only certain types of cells, making their presence or absence helpful in 
classifying cells. 

Two of the most common CDs to include on a panel are CD3, CD4, and CD8. T cells have
CD3 on their surface, so a marker for CD3 can be used to distinguish T cells
from other types of white blood cells, including granulocytes, monocytes, and
B cells. Among T cells, the helper T cells have the CD4 protein on their surface, 
while the cytotoxic T cells have the CD8 protein on their surface, so the CD4 
and CD8 markers can help in refining a T cell into a more specific type. 

In flow cytometry, you can characterize immune cells into populations based on
proteins on the cell surface and inside the cell, as well as cell size and
granularity [@maecker2012standardizing, @barnett2008cd4]. For example,
macrophages can be distinguished from T cells and B cells based on ... [size?
granularity?], which T cells and B cells can be distinguished from each other
based on whether the cell has the [surface protein? CD3?], and helper T cells
versus cytotoxic T cells can be distinguished from each other based on whether
the cell has the [surface protein CD8? CD4?].

## Loading packages

```{r, warning=FALSE, echo=FALSE, message = FALSE}
library(readxl)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(tidyverse)
library(scales)
library(stringr)
library(tidyr)
library(knitr)
library(forcats)
library(broom)
library(ggfortify)
library(stats)
library(ggpubr)
library(grDevices)
library(rstatix)
library(writexl)
library(purrr)
library(kableExtra)
library(ggbeeswarm)
library(multcomp)
library(DescTools)
library(car)
library(exactRankTests)
library(readr)

```
MULTI CSV SHEETS MALE FEMALE

## Loading data

```{r}
# Read in all CSV files
files <- list.files(path = "DATA/flow data", pattern = "*.csv", full.names = TRUE)

dfs <- lapply(files, function(file) {
  # Read the CSV file and clean up column names
  df <- read_csv(file)
    df <- df %>%
    dplyr::select(-matches("Parent")) %>%
    rename_all(~ str_remove(., "Leukocytes/Single Cells/LIVE/")) %>%
    rename_all(funs(str_replace(., "\\|.+", ""))) %>%
    dplyr::select(-matches("Leukocytes")) %>%
    rename("Sample" = colnames(.)[1]) %>%
    mutate(Sample = str_replace_all(Sample, "\\b[A-Za-z][0-9]+\\b", "")) %>%
    rename_all(funs(str_replace_all(., ",", ""))) %>%
    rename_all(funs(str_replace_all(., "/", " "))) %>%
    rename_all(funs(str_replace_all(., "\\\\", " ")))  %>%
    rename_all(~ stringr::str_replace_all(., "Q\\d+\\:", "")) %>%
  # Removing logical vectors, not sure why its adding these
    dplyr::select(where(~ !any(is.logical(.)) | is.numeric(.))) %>%
  # Remove last two rows of SD and Mean
  slice(1:(n() - 2))
  # Extract the sex from the file name
  sex <- ifelse(grepl("female", file), "female", "male")
  df <- df %>% mutate(sex = sex)
  # Move the sex column to the second position
  df <- df %>% relocate(sex, .after = 1)
  # Extract the day from the file name
  day <- str_extract(file, "(?<=Day\\s)\\d+")
  df <- df %>% mutate(day = as.character(day))
  # Move the day column to the third position
  df <- df %>% relocate(day, .after = 2)
  
  return(df)
})

# Combine all data frames into one dataframe
Df1 <- bind_rows(dfs)
Df1

#str(dfs)
#tail(Df1)

# Check for NA values in Df1
na_counts <- colSums(is.na(Df1))

# Print the column names with NA counts
print(na_counts)

```
MULTIDAY SHEETS FOR EXCEL
```{r}
# # Read in all sheets from the Excel file MULTIDAY
# sheets <- excel_sheets("/Volumes/cvmbs/RSTOR-HenaO_lab/SolaVAX-TB/SolaVax Prime/Day 3/Myeloid Lung D3 populations.xlsx")
# 
# dfs <- lapply(sheets, function(sheet) {
# # Read in the sheet and clean up column names
#   df <- read_excel("/Volumes/cvmbs/RSTOR-HenaO_lab/SolaVAX-TB/SolaVax Prime/Day 3/Myeloid Lung D3 populations.xlsx", sheet = sheet) %>%
#     dplyr::select(-matches("Parent")) %>%
#     rename_all(~ str_remove(., "Leukocytes/Single Cells/LIVE/")) %>%
#     rename_all(funs(str_replace(., "\\|.+", ""))) %>%
#     dplyr::select(-matches("Leukocytes")) %>%
#     dplyr::select(-2) %>%
#     rename("Sample" = "Sample:") %>%
#     mutate(Sample = str_replace_all(Sample, "\\b[A-Za-z][0-9]+\\b", "")) %>%
#     rename_all(funs(str_replace_all(., "\\/Q[[:digit:]]+\\:", ""))) %>%
#     rename_all(funs(str_replace_all(., ",", ""))) %>%
#     rename_all(funs(str_replace_all(., "/", " "))) %>%
#     rename_all(funs(str_replace_all(., "\\\\", " ")))
# 
# # Add the Day column
#   day <- str_extract(sheet, "(?<=Day\\s)\\d+")
#   df <- df %>% mutate(Day = as.character(day))
# 
# # Move the Day column to the second position
#   df <- df %>% relocate(Day, .after = 1)
# 
#   return(df)
# })
# 
# # Combine all sheets into one data frame
# Df1 <- bind_rows(dfs)
# Df1

```

SINGLE SHEETS Loading data
```{r eval=FALSE, include=FALSE}
# Df<- read_csv("/Users/pablo/Desktop/IMPAcTB/male.csv")
# 
# # Remove Freq of Parent columns
# Df1 <- Df %>%
#   dplyr::select(-matches("Parent"))
# # Remove "Leukocytes/LIVE/Single Cells/" from col names
# names(Df1) <- str_remove(names(Df1), "Leukocytes/Single Cells/LIVE/")
# 
# Df1 <- Df1 %>%
#   rename_all(funs(str_replace(., "\\|.+", "")))# Remove "|Freq of..." from col names
# 
# 
# Df1 <- Df1 %>%
#   dplyr::select(-matches("Leukocytes")) %>%
#   dplyr::select(-2)
# 
# # Remove the colon from the column name
# names(Df1)[names(Df1) == "Sample:"] <- "Sample"
# 
# # Rename the column values by removing the first two letters and numbers
# Df1 <- Df1 %>%
#   mutate(Sample = str_replace_all(Sample, "\\b[A-Za-z][0-9]+\\b", ""))
# 
# 
# Df1 <- Df1 %>%
#    rename_all(funs(str_replace_all(., ",", ""))) %>%
#    rename_all(funs(str_replace_all(., "/", " "))) %>%
#    rename_all(funs(str_replace_all(., "\\\\", " ")))  %>%
#   rename_all(~ stringr::str_replace_all(., "Q\\d+\\:", "")) 
# 
# Df1 <- Df1 %>% slice(1:(n() - 2))
# 
# 
# #Not needed for this 
# # Df1 <- Df1 %>%
# #   mutate(Day = 1)
```


MAKING DATA TIDY FOR PLOTTING
```{r}
tidy_Df1 <- pivot_longer(data = Df1, cols =  starts_with("CD3"), names_to = "cell_types", values_to = "percentage_of_LIVE")

tidy_Df1 <- tidy_Df1 %>%
  separate(col = "Sample", into = c("group", "mouse_ID"), sep = "_") %>%
  mutate(mouse_ID = str_replace(mouse_ID, ".fcs", ""))

#Trim extra spaces
tidy_Df1$cell_types <- str_replace_all(tidy_Df1$cell_types, "\\s+", " ")
tidy_Df1$cell_types <- str_trim(tidy_Df1$cell_types)
tidy_Df1$group <- str_replace_all(tidy_Df1$group, "\\s+", " ")
tidy_Df1$group <- str_trim(tidy_Df1$group)

tidy_Df1

str(tidy_Df1)

tidy_Df1$group <- str_replace_all(tidy_Df1$group, "-", " ")
```

CHECK FIRST ASSUMPTION IN KRUSKAL (Independent obervations)
```{r}
# We want to plot one population and facet by day, fill by sex

# Get unique values in "cell_types" column to check for spelling
unique(tidy_Df1$cell_types)
# I am choosing "CD3+ CD4+ CD62L- CD44+"

test_pop_assumption1 <- tidy_Df1 %>% 
  dplyr::filter(cell_types == "CD3+ CD4+ CD62L- CD44+")

#plot
ggplot(test_pop_assumption1, aes(x = group, y = percentage_of_LIVE, color = sex)) +
  geom_dotplot(binaxis = "y", stackdir = "center", fill = "white", dotsize = 0.5) +
  facet_wrap(~ day, ncol = 1) +
  theme_bw() +
  labs(x = "Sex", y = "CD4+ Effector TCells (% of Live)", color = "Sex") +
  ggtitle("Dotplot Distribution of Cell Types by Sex and Day")



#how to check if i need to do one factor or two factor anova
```


WORKING ON STATS June 1, 2023
```{r}
# library(emmeans)
# library(car)
# library(DescTools)
# library(stats)
# 
# 
# tidy_Df2 <- tidy_Df1 %>% 
#   dplyr::filter(cell_types == " CD3-  CD11B+ ")
# 
# 
# str(tidy_Df2)
# 
# CD3CD11b_LM<- lm(percentage_of_LIVE ~ group, data = tidy_Df2)
# 
# plot(CD3CD11b_LM)
# 
# 
# library(tidyverse)
# library(dplyr)
# library(kableExtra)
# 
# sumstats<-tidy_Df2%>%group_by(group)%>%summarise(n=n(),mean=mean(percentage_of_LIVE,na.rm=TRUE),sd=sd(percentage_of_LIVE,na.rm=TRUE),sem=sd(percentage_of_LIVE,na.rm=TRUE)/sqrt(n),min=min(percentage_of_LIVE,na.rm=TRUE),max=max(percentage_of_LIVE,na.rm=TRUE))
# 
# 
# kable(sumstats,digits=2)%>%kable_styling(full_width = F)
# 
# sumstats
# 
# #Krsukal check then do tukey or duns test for multiple comparispons need multcomp 
# 
# library(car)
# library(emmeans)
# library(DescTools)
# library(car)
# library(emmeans)
# library(DescTools)
# 
# ANOVA <- Anova(lm(percentage_of_LIVE~group, data = tidy_Df2),type='II')
# 
# ANOVA
# 
# 
# ANOVA.emm <- emmeans(CD3CD11b_LM,~ group)
# 
# pairs(ANOVA.emm,adjust="tukey",type="response")
# 
# 
# bartlett.test(percentage_of_LIVE ~ group, data = tidy_Df2)
```

EXPLORATORY ANALYSIS
```{r}

# Running Stats and Selecting Significant Cell Type Differences for Gated Cells TukeyHSD
cell_stats <- tidy_Df1 %>%
   group_by(cell_types) %>%
  nest() %>%
  mutate(aov_result = map(data, ~aov(percentage_of_LIVE ~ group, data = .x)),
         tukey_result = map(aov_result, TukeyHSD),
         tidy_tukey = map(tukey_result, broom::tidy)) %>%
  unnest(tidy_tukey, .drop = TRUE) %>%
  separate(contrast, into = c("contrast1", "contrast2"), sep = "-") %>%
  dplyr::select(-data, -aov_result, -tukey_result, -term, -null.value) %>%
  filter(adj.p.value <= 0.05)


# Join Dataframes Based on Significance and Cell Types
common_cell_types <- intersect(tidy_Df1$cell_types, cell_stats$cell_types)
tidy_Df1_filtered <- tidy_Df1 %>%
  filter(cell_types %in% common_cell_types)

joined_df <- left_join(tidy_Df1_filtered, cell_stats, by = "cell_types")

unique(joined_df$group)



```


PLOT POPULATIONS
```{r}
# Get the Dark2 palette
my_palette <- brewer.pal(n = 8, name = "Dark2")

# Choose specific colors from the palette
my_colors <- c(my_palette[1], my_palette[3], my_palette[5], my_palette[8], my_palette[2])


# Multiday Plot
joined_df$group <- factor(joined_df$group, levels = c("Saline", "BCG", "ID93", "BCG+ID93"))

ggplot(joined_df, aes(x = group, y = percentage_of_LIVE, fill = group)) +
geom_boxplot(aes(x = group, y = percentage_of_LIVE), linetype = 1, width = .5, alpha = 0.5) + 
geom_point(show.legend = FALSE)+
 theme_bw() + 
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10, face = "bold"),
        axis.title.y = element_text(size = 10, face = "bold"),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        plot.title = element_text(color="black", size=10, face="bold")) +
  labs(x = "Vaccine", y = "Percentage of Live Leukocytes", color = "Group")+
  guides(shape = "none") +
  theme(strip.text = element_text(size=8, face = "bold")) + 
  theme(legend.position="right") +
ggtitle("IMPAcTB Males Females Combined ") +
  facet_wrap(~reorder(cell_types, -percentage_of_LIVE) + day, 
             scales = "free_y", labeller = label_wrap_gen(width=20), 
             ncol = 4, nrow = 50, strip.position = "top") +
  labs(fill = "Vaccine") 

```

WORKING ON THIS 
```{r}
#+
   #scale_fill_manual(values = my_colors)

# Adding p-values from tukeyHSD Dataframe 
# p + geom_text(data = joined_df %>%ast 
#                 group_by(contrast1, contrast2) %>%
#               aes(label = paste0("p = ", round(adj.p.value, 4)),
#                   y = percentage_of_CD45 + 0.001),
#               size = 6.5, hjust = "center", vjust = 20) 


# Adding p-values from tukeyHSD Dataframe 
# Interest_Plots <- p1 + geom_text(data = cell_types_of_interest %>%
#                 group_by(cell_types) %>%
#                 slice_max(percentage_of_CD45),
#               aes(label = paste0("p = ", round(adj.p.value, 4)),
#                   y = percentage_of_CD45 + 0.001),
#               size = 6.5, hjust = "center", vjust = 20) +
#   scale_fill_manual(values = my_colors)
#   
# Interest_Plots
```

Saving plots to PDF
```{r}
# #save to pdf
# pdf(file = "/Volumes/cvmbs/RSTOR-HenaO_lab/Pablo/SOLOVAX TB/SolovaxTB immunogenicity day 14/Unmixed/FlowJo/CD45_Final_Lung_Plots.pdf",   # The directory you want to save the file in
#    width = 25, # The width of the plot in inches
#    height = 35) # The height of the plot in inches
# plot(CD45_Final_Lung_Plots)
# dev.off()
```


Looking at specific populations
```{r fig.height=2, fig.width=3}

#Finding correct spelling and punctuation of cell_types EXAMPLE

tidy_Df1$cell_types [23]

#Showing Neutrophil Populations

CD11b_Mac <- tidy_Df1 %>% 
  filter(cell_types == "CD45 Ly6G- CD64- CD11c-  CD11b+ ")

Ly6G_df <- tidy_Df1 %>% 
  filter(cell_types == "CD45 Ly6G+ ")

Ly6G_stats <- Ly6G_df %>%
   group_by(cell_types) %>%
  nest() %>%
  mutate(aov_result = map(data, ~aov(percentage_of_CD45 ~ group, data = .x)),
         tukey_result = map(aov_result, TukeyHSD),
         tidy_tukey = map(tukey_result, broom::tidy)) %>%
  unnest(tidy_tukey, .drop = TRUE) %>%
  separate(contrast, into = c("contrast1", "contrast2"), sep = "-") %>%
  dplyr::select(-data, -aov_result, -tukey_result, -term, -null.value)

common_cell_types <- intersect(Ly6G_df$cell_types, Ly6G_stats$cell_types)
Ly6G_df_filtered <- Ly6G_df %>%
  filter(cell_types %in% common_cell_types)

Ly6G_joined_df <- left_join(Ly6G_df_filtered, Ly6G_stats, by = "cell_types")


ggplot(Ly6G_joined_df, aes(x = group, y = percentage_of_CD45, fill = group, alpha=0.5)) +
  geom_boxplot() + 
  geom_beeswarm() + 
  facet_wrap(~cell_types, scales = "free_y", labeller = label_wrap_gen(width=20), ncol = 6, nrow = 50) +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.text.y = element_text(size = 8),
        axis.title.x = element_text(size = 10, face = "bold"),
        axis.title.y = element_text(size = 10, face = "bold"),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        plot.title = element_text(color="black", size=10, face="bold")) +
  labs(x = "Group", y = "Percentage of Live Leukocytes") +
  theme(strip.text = element_text(size=8, face = "bold")) + theme(legend.position="right") +
  ggtitle("Neutrophil Response 14 Days Post Vaccination
C57Bl/6 mice (Immunogenicty Study-Non-Infected)") +
  geom_text(data = Ly6G_joined_df %>% 
                group_by(cell_types) %>% 
                slice_max(percentage_of_CD45),
              aes(label = paste0("p = ", round(adj.p.value, 4)), 
                  y = percentage_of_CD45 + 0.01), 
              size = 3, hjust = "right") +
  scale_fill_manual(values = my_colors)
```


```{r fig.height=10, fig.width=15}
# Checking for correct spelling 
unique(joined_df$cell_types)

# New DF with cell types of interest (will always change)
cell_types_of_interest <- joined_df %>%
  filter(cell_types %in% c("CD45 Ly6G- CD64- CD11c-  CD11b- Ly6C- ", 
                           "CD45 Ly6G- CD64+ ", 
                           "CD45 Ly6G- CD64- CD11c-  CD11b+ Ly6C+ ", 
                           "CD45 Ly6G- CD64- CD11c-  CD11b+ Ly6C+ CD26-  CD88- ", 
                           "CD45 Ly6G- CD64+ CD11c+  CD11b- ", 
                           "CD45 Ly6G- CD64+ CD11c+  CD11b+ ",
                           "CD45 Ly6G- CD64+ CD11c+  CD11b- Ly6C- ",
                           "CD45 Ly6G- CD64- CD11c+  CD11b+ Ly6C+ ",
                           "CD45 Ly6G- CD64+ CD11c+  CD11b- Ly6C- SiglecF+  CD206+ ",
                           "CD45 Ly6G- CD64+ CD11c+  CD11b+ Ly6C- ",
                           "CD45 Ly6G- CD64+ CD11c+  CD11b+ Ly6C- CD86+  MHCII+ ",
                           "CD45 Ly6G- CD64+ CD11c+  CD11b+ Ly6C+ CD86+  MHCII+ ",
                           "CD45 Ly6G- CD64- CD11c+  CD11b+ Ly6C+ CD86+  MHCII+ ",
                           "CD45 Ly6G- CD64+ CD11c-  CD11b+ Ly6C+ CD86+  MHCII+ ",
                           "CD45 Ly6G- CD64+ CD11c+  CD11b+ Ly6C- SiglecF-  CD206- "))


# Plot again
p1 <- ggplot(cell_types_of_interest, aes(x = group, y = percentage_of_CD45, fill = group, alpha=0.5)) +
  geom_boxplot() + 
  geom_beeswarm(groupOnX = FALSE) + 
  facet_wrap(~cell_types, scales = "free_y", labeller = label_wrap_gen(width=30), ncol = 5, nrow = 50) +
  theme_bw() + 
  theme(axis.text.x = element_blank(), axis.text.y = element_text(size = 20),
        axis.title.x = element_text(size = 30, face = "bold"),
        axis.title.y = element_text(size = 30, face = "bold"),
        legend.text = element_text(size = 30),
        legend.title = element_text(size = 30),
        plot.title = element_text(color="black", size=30, face="bold")) +
  labs(x = "Group", y = "Percentage of Live Leukocytes") +
  theme(strip.text = element_text(size=18, face = "bold")) + theme(legend.position="right") +
  ggtitle("SolaVax TB Myeloid Response 14 Days Post Vaccination in C57Bl/6 mice (Immunogenicty Study-Non-Infected)") +
   facet_wrap(~reorder(cell_types, -percentage_of_CD45), 
             scales = "free_y", labeller = label_wrap_gen(width=20), 
             ncol = 5, nrow = 50, strip.position = "top")

```
