---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Proteomics

For proteomics data, we will be getting data that have already been collected and 
pre-processed by another part of the team. The following shows an example of the 
type of data we will get as an input:

```{r}
library(tidyverse)

prot_a <- read_csv("DATA/Transition Results_CCTSI_A.csv")

prot_a
```

These data include the following columns: 

- `Peptide`: A short string of peptides that are being measured
- `Protein`: The protein that those peptides come from
- `Replicate`: An identifier for the sample that the measurement was taken on
- `Precursor Mz`, `Precursor Charge`, `Product Mz`, `Product Charge`, 
`Fragment Ion`, `Retention Time`: Measurements that help in identifying the peptide
that is being measured (?)
- `Area`: 
- `Background`: 
- `Peak Rank`: 
- `Ratio Dot Product`: 
- `Total Area Normalized`: 
- `Total Area Ratio`
- `Library Dot Product`: 
- `RatioLightToHeavy`: 
- `DotProductLightToHeavy`: 

[More about how these data were pre-processed. Softwarei: Skyline]

Here are all the unique replicates in this file: 

```{r}
prot_a %>% 
  pull(Replicate) %>% 
  unique()
```

The three groups in this data are labeled with "LT", "H", and "TB" somewhere in 
the identifier. We can create a new column in the dataset that pulls out this
treatment group information: 

```{r}
prot_a <- prot_a %>% 
  mutate(treatment_group = str_extract(Replicate, "[A-Z]+")) 

prot_a %>% 
  filter(Peptide == first(Peptide)) %>% 
  group_by(treatment_group) %>% 
  count()

prot_a %>% 
  filter(Peptide == first(Peptide) & 
           Replicate == first(Replicate))
``` 


```{r}
prot_a %>% 
  pull(Protein) %>% 
  unique()
```

- Cfp10
- acpM
- Ag85A
- MtbH37Rv|Rv3841|BfrB
- MtbH37Rv|Rv1837c|GlcB
- MtbH37Rv|Rv3418c|GroES
- MtbH37Rv|Rv3248c|SahH
- MtbH37Rv|Rv2031c|hspX
