--- 
title: "ImpactTB/BAA: Standard Operating Procedures for Data Analysis"
author: "Colorado State University Coding Team"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "A comprehensive guide to wet lab data collection, sample processing, and computational tool creation for robust and efficient data analysis and dissemination."
---
--- 
title: "ImpactTB/BAA: Standard Operating Procedures for Data Analysis"
author: "Colorado State University Coding Team"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "A comprehensive guide to wet lab data collection, sample processing, and computational tool creation for robust and efficient data analysis and dissemination."
---

# Overview

Here, we have built a comprehensive guide to wet lab data collection, sample processing, and computational tool creation for robust and efficient data analysis and dissemination.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

<!--chapter:end:index.Rmd-->

# Colony forming units to determine bacterial counts TESTING !!!!

Selecting dilution for CFU {#CFU}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

<!--chapter:end:01-intro.Rmd-->

# Initial mouse characteristics

Here is a review of existing methods.

<!--chapter:end:02-initial_mouse_characteristics.Rmd-->

# Mouse weights


Mice are weighed in grams weekly and recorded in an excel worksheet. Column titles are as follows: who_collected	date_collected	sex	dob	notch_id	mouse_number	weight	unit	cage_number	group	notes



```{r}
library(readxl)
library(tidyverse)
```

## Read in data
```{r}
weight_data <- read_xlsx()
```


<!--chapter:end:03-mouse_weights.Rmd-->

# Colony forming units to determine bacterial counts 

## Data description

The data are collected in a spreadsheet with multiple sheets. The first sheet
(named "[x]") is used to record some metadata for the experiment, while the 
following sheets are used to record CFUs counts from the plates used for samples
from each organ, with one sheet per organ. For example, if you plated data
from both the lung and spleen, there would be three sheets in the file: one 
with the metadata, one with the plate counts for the lung, and one with the
plate counts for the spleen. 

The metadata sheet is used to record information about the overall process of
plating the data. Values from this sheet will be used in calculating the bacterial
load in the original sample based on the CFU counts. This spreadsheet includes
the following columns: 

- `organ`: Include one row for each organ that was plated in the experiment. 
You should name the organ all in lowercase (e.g., "lung", "spleen"). You 
should use the same name to also name the sheet that records data for that organ
for example, if you have rows in the metadata sheet for "lung" and "spleen", 
then you should have two other sheets in the file, one sheet named "lung" and 
one named "spleen", which you'll use to store the plate counts for each of those
organs.
- `prop_resuspended`: In this column, give the proportion of that organ that 
was plated. For example, if you plated half the lung, then in the "lung" row
of this spread sheet, you should put 0.5 in the `prop_resuspended` column. 
- `total_resuspended_uL`: This column contains an original volume of tissue homogenate. For example, raw lung tissue is homogenized in 500 uL of PBS in a tube containing metal beads. 
- `og_aliquot_uL`: 100 uL of th total_resuspended slurry would be considered an original aliquot and is used to peform serial dilutions.
- `dilution_factor`: Amount of the original stock solution that is present in the total solution, after dilution(s)
- `plated_uL`: Amount of suspension + diluent plated on section of solid agar 

## Read in data

```{r message = FALSE, warning = FALSE}
library(readxl)
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)

#Replace w/ path to CFU sheet
path <- c("~/cfu_prototype/example_spreadsheet/Copy of baa_cfu_sheet.xlsx")

sheet_names <- excel_sheets(path)
sheet_names <- sheet_names[!sheet_names %in% c("metadata")]

merged_data <- list()

for(i in 1:length(sheet_names)){
  
  data <- read_excel(path, sheet = sheet_names[i]) %>% 
    mutate(organ = paste0(sheet_names[i]))
  
  data <- data %>% 
    #mutate(missing_col = NA) %>% 
    mutate_if(is.double, as.numeric) %>% 
    mutate_if(is.numeric, as.character) %>% 
    pivot_longer(starts_with("dil_"), names_to = "dilution",
                 values_to = "CFUs") %>% 
    mutate(dilution = str_extract(dilution, "[0-9]+"),
           dilution = as.numeric(dilution))
    
  
  merged_data[[i]] <- data
  
  
}
  
all_data <- bind_rows(merged_data, .id = "column_label") %>% 
    select(-column_label)
  
```

## Exploratory analysis and quality checks

### Exploratory analysis

**Dimensions of input data:**

Based on the input data, data were collected for the following organ or 
organs: 

The following number of mice were included for each: 

The following number of replicates were recorded at each count date for 
each experimental group: 

The following number of dilutions and dilution level were recorded for 
each organ: 

**People who plated and collected the data. Date or dates of counting:**

Based on the input data, the plates included in these data were counted by 
the following person or persons: 
Based on the input data, the plates included in these data were counted on 
the following date or dates: 

```{r}
all_data %>%
  select(organ, who_plated, who_counted, count_date) %>%
  distinct()
```

**Distribution of CFUs at each dilution:**

WE NEED TO ADD SAMPLE CFU PLOTS

Here's a plot that shows how many plates were too numerous to count at each 
dilution level: 

Here is a plot that shows how the CFU counts were distributed by dilution
level in the data: 

## Identify a good dilution for each sample
```{r}
# Make all_data into tidy data and filter for CFUs between 10-75
  
tidy_cfu_data <- all_data %>%
  mutate(dilution = str_extract(dilution, "[0-9]+"),
         dilution = as.numeric(dilution)) %>%
  filter(CFUs >= 10 & CFUs <= 75) %>%
  mutate(CFUs = as.numeric(CFUs))
```

## Calculate CFUs from best dilution/Estimate bacterial load for each sample based on good dilution
```{r}
# Calculating CFU/ml for every qualifying replicate between 10-75 CFUs. Column binding by organ name to the metadata sheet via inner_join().
meta <- read_excel(path, sheet = "metadata")

tidy_cfu_meta_joined <- inner_join(tidy_cfu_data, meta) %>%
  group_by(groups) %>% 
  mutate(CFUs_per_ml = (CFUs * (dilution_factor^2) * (total_resuspension_mL/volume_plated_ul) * 10)) %>%
  select(organ, count_date, who_plated, who_counted, groups,  mouse, dilution,  CFUs, CFUs_per_ml) %>%
  ungroup()

tidy_cfu_meta_joined
```

## Create initial report information for these data
```{r}

```


## Sample ANOVA
```{r}
cfu_stats <- tidy_cfu_meta_joined %>% 
  group_by(organ) %>%
  nest() %>%
  mutate(aov_result = map(data, ~aov(CFUs_per_ml ~ groups, data = .x)),
         tukey_result = map(aov_result, TukeyHSD),
         tidy_tukey = map(tukey_result, broom::tidy)) %>%
  unnest(tidy_tukey, .drop = TRUE) %>%
  separate(contrast, into = c("contrast1", "contrast2"), sep = "-") %>%
  select(-data, -aov_result, -tukey_result, -term, -null.value)# %>%
  # filter(adj.p.value <= 0.05)

cfu_stats
```

## Save processed data to database
## Example one

## Example two

<!--chapter:end:04-cfus.Rmd-->

# ELISA Words

We have finished a nice book.

<!--chapter:end:05-elisa.Rmd-->

`r if (knitr:::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

